---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tdid

<!-- badges: start -->
<!-- badges: end -->

These R functions calculate triple difference-in-differences (TDID) with controls using double-robust estimators. 

## Installation

You can install the development version of tdid from [GitHub](https://github.com/) with:

``` {r}
devtools::install_github("dorleventer/tdid")
```

## Example

Load a sample based on the DGP of the example in the paper. 

```{r}
library(tdid)
library(tidyverse)
library(fixest)
data = example_dgp(n = 1000, seed = 1)
head(data)
```
The causal estimand is equal to three. 

Running a three-way linear specification with controls.
```{r}
data = data |>
  mutate(P = 1*(time == 2),
         A = 1*(group == "A"))
tdid_linear = coef(lm(Y ~ W*P*A + X, data = data))[9]
cat("Estimated diff in ATT via linear specification: ",round(tdid_linear,2))
```

A difference in DR estimates by group

```{r}
dr_A = DRDID::drdid(yname = "Y", tname = "time", idname = "id", dname = "W", xformla = ~X, data = data |> filter(group == "A"))$ATT
dr_B = DRDID::drdid(yname = "Y", tname = "time", idname = "id", dname = "W", xformla = ~X, data = data |> filter(group == "B"))$ATT
diff = as.numeric(dr_A - dr_B)
cat("Estimated diff in ATT via difference in DR estimators within groups: ",round(diff,2))
```

The DR TDID estimator

```{r}
tdid_dr = drtdid(data)$tau
cat("Estimated diff in ATT via difference in DR of group A and reweighted DR of group B: ",round(tdid_dr,2))
```

