---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tdid

<!-- badges: start -->
<!-- badges: end -->

These R functions calculate triple difference-in-differences (TDID) with controls using double-robust estimators. 

## Installation

You can install the development version of tdid from [GitHub](https://github.com/) with:

``` {r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
devtools::install_github("dorleventer/tdid")
```

## Example

An example, as well as replication of the graph from the paper, is given in the [example vignette]( https://dorleventer.github.io/tdid/articles/tdid-example.html).

## In Progress

Currently this is a bare-bones package repo for replication of the results in the paper. In the future will upgrade such that

- [ ]  Handle different names from outcome, time, ...
- [ ]  Allow inputting covariate formulas
- [ ]  Useage of different ML learners for the propensity score and outcome regression
- [ ]  Cross-fitting of the propensity score and outcome regression
