---
title: "tdid-application"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tdid-application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Replication

First, for completness, here is the R code from the mixtape, where Scott discusses the TDID design. (Note that i made minor changes)

```{r setup}

library(tidyverse)
library(haven)
library(estimatr)

read_data <- function(df)
{
  full_path <- paste("https://github.com/scunning1975/mixtape/raw/master/", 
                     df, sep = "")
  df <- read_dta(full_path)
  return(df)
}

abortion <- read_data("abortion.dta") %>% 
  mutate(
    repeal  = as_factor(repeal),
    year    = as_factor(year),
    fip     = as_factor(fip),
    fa      = as_factor(fa),
    younger = as_factor(younger),
    yr      = as_factor(case_when(repeal == 1 & younger == 1 ~ 1, TRUE ~ 0)),
    wm      = as_factor(case_when(wht == 1 & male == 1 ~ 1, TRUE ~ 0)),
    wf      = as_factor(case_when(wht == 1 & male == 0 ~ 1, TRUE ~ 0)),
    bm      = as_factor(case_when(wht == 0 & male == 1 ~ 1, TRUE ~ 0)),
    bf      = as_factor(case_when(wht == 0 & male == 0 ~ 1, TRUE ~ 0))
  ) %>% 
  filter(bf == 1 & (age == 15 | age == 25))

regddd <- lm_robust(lnr ~ younger*repeal*year + acc + pi + alcohol + crack + poverty + income + ur,
                    data = abortion, weights = totpop, clusters = fip)

abortion_plot <- tibble(
  sd = regddd$std.error[58:72],
  mean = regddd$coefficients[58:72],
  year = c(1986:2000))

abortion_plot %>% 
  ggplot(aes(x = year, y = mean)) + 
  geom_point()+
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = mean-sd*1.96, ymax = mean+sd*1.96), width = 0.2,
                position = position_dodge(0.05)) + 
  ylim(-1, 1) + 
  scale_x_continuous(breaks = 1986:1995, limits = c(1985.5, 1995.5))
```

# The design

First, lets use the notation in the paper for this application. 

- Time periods ($t$) are years.
- States which repeal abortion early ($E=1986$) vs late ($E=1989$). 
- Groups are young ($G_i=15-19$) and old ($G_i=25-29$). 

# DID with controls estimators by group

This is the data for the analysis

```{r}
abortion <- read_data("abortion.dta") %>% 
  mutate(
    repeal  = as.numeric(repeal),
    year    = as.numeric(year),
    younger = as.numeric(younger),
    bf      = as.numeric(case_when(wht == 0 & male == 0 ~ 1, TRUE ~ 0))
  ) %>% 
  filter(bf == 1 & (age == 15 | age == 25))
```

Lets rename the variables into our notation

```{r}
data_analysis <- abortion |> 
  select(Y = lnr, 
         id, totpop, fip,
         W = repeal,
         time = year, 
         group = younger,
         acc, pi, alcohol, crack, poverty, income, ur)
```

And run DID, with 1993 as the reference year, separately by group. 

First, lets code a data preparation function, to make a 2-by-2 comparison

```{r}
# data_prep = function(data, target_year, reference_year, control_formula) {
#   
#   # drop NAs
#   data = data |> tidyr::drop_na()
#   
#   # make panel
#   time_vec  <- data[["time"]]
#   data = data[which(time_vec == target_year | time_vec == reference_year),]
#   data = BMisc::makeBalancedPanel(data, "id", "time")
#   
#   # construct a 2-by-2 data frame
#   time_vec      <- data[["time"]]
#   id_vec        <- data[["id"]]
#   y_vec         <- data[["Y"]]
#   treat_vec     <- data[["W"]]
#   weight_vec    <- data[["totpop"]]
#   cluster_vec   <- data[["fip"]]
#   group_var     <- data[["group"]]
# 
#   ids      <- id_vec[time_vec == target_year]
#   y1       <- y_vec[time_vec == reference_year]
#   y2       <- y_vec[time_vec == target_year]
#   y_diff   <- y2 - y1
#   W        <- treat_vec[time_vec == target_year]
#   weights  <- weight_vec[time_vec == target_year]
#   clusters <- cluster_vec[time_vec == target_year]
#   groups   <- group_var[time_vec == target_year]
#   
#   # construct covariate matrix
#   if (inherits(control_formula, "formula")) {
#     formula_obj <- control_formula
#   } else {
#     formula_obj <- as.formula(paste0("~", control_formula))
#   }
# 
#   X <- stats::model.matrix(
#     formula_obj,
#     stats::model.frame(formula_obj, data = data[time_vec == target_year, ], na.action = na.pass)
#   )
# 
#   return(list(
#     data_did = data.frame(ids = ids, y1 = y1, y2 = y2, y_diff = y_diff, W = W, weights = weights, clusters = clusters, groups = groups),
#     X_mat = X[, -1, drop = FALSE]
#   ))
# }
```

Next, lets walk through the double robust approach

Consider the object 
$$\mathbb{E}\big[\mathbb{E}[Y_{t}\mid G=g^\prime,E=e^\prime,X]\mid G=g,E=e\big]$$
This is equals to 
$$\mathbb{E}\big[w_t(g,e)\mathbb{E}[Y_{t}\mid G=g^\prime,E=e^\prime,X]\big],\quad w_t(g,e)=\frac{1_{G_i=g,E_i=e}}{\mathbb{E}[1_{G_i=g,E_i=e}]}$$

Next, we can write the outcome regression (OR) identification formula, like so 

$$\mathbb{E}\big[w_{\mathrm{T}}(g,e)\mathbb{E}[Y_{t}\mid G=g^\prime,E=e^\prime,X]\big]=\mathbb{E}\big[w_t(g,e)m(g^\prime,e^\prime,t,X)\big],\quad m(g^\prime,e^\prime,t,X)=\mathbb{E}[Y_{t}\mid G=g^\prime,E=e^\prime,X=x]$$

And the inverse probability weighting (IPW) identification formula, like so 

$$\mathbb{E}\big[w_{\mathrm{T}}(g,e)\mathbb{E}[Y_{t}\mid G=g^\prime,E=e^\prime,X]\big]=\mathbb{E}\big[w_{\mathrm{C}}(g,g^\prime,e,e^\prime,X)Y_t\big],\quad w_{\mathrm{C}}(g,g^\prime,e,e^\prime,x)=\frac{1_{G_i=g^\prime,E_i=e^\prime}}{\mathbb{E}[1_{G_i=g,E_i=e}]}\frac{\mathbb{E}[1_{G_i=g,E_i=e}\mid X=x]}{\mathbb{E}[1_{G_i=g^\prime,E_i=e^\prime}\mid X=x]}$$

When thinking about using these formulas in our context, we need to note the following. 

```{r}
# For example, consider the target year 1986, using 1993 as reference
# prep = data_prep(data = data_analysis, target_year = 1986, reference_year = 1993, control_formula = "acc + pi + alcohol + crack + poverty + income + ur")
# prep$data_did |> group_by(groups, W) |> count()
```

As we can see, there are 5 observations with $W=1$. 

- Note that the covariate vector in the regression was acc + pi + alcohol + crack + poverty + income + ur 
- That is, there are 7 covariates, and 5 observations
- Below we use only 4 of these covariates, but still, very low
- When doing OR in TDID, when of the regressions will use units with $W=1$ and $W=0$ (older black females in repeal states)
- But we can't really run such a regression, so will use only IPW below (which uses all observations when calculating propensity scores)

Lets code some functions to estimate logistic regressions for propensity scores

```{r}
# train_log = function(Y, X, weights) {
#   glm(Y ~ ., family = binomial(), data = data.frame(Y, X), weights = weights)
# }
# pred_log = function(mod, X) {
#   predict(mod, data.frame(X), type = "response")
# }
# learning_function <- function(Y, X, ids_out, ids_in, weights) {
# 
#   # For the non-crossfitting branch, only use the subset of observations in ids_in for training.
#   train_idx <- which(ids_out %in% ids_in)
#   mod <- train_log(Y[train_idx], X[train_idx, , drop = FALSE], weights[train_idx])
#   res <- as.numeric(pred_log(mod, X))
#   
#   return(res)
# }
```

This is a TDID hand written function, based on IPW identification

```{r}
# tdid_function = function(data_analysis, target_year, reference_year, control_formula) {
#   prep = data_prep(data = data_analysis, target_year, reference_year, control_formula)
#   
#   data_did <- prep$data_did
#   ids <- data_did$ids
#   
#   indt_1 <- as.numeric(data_did$W == 1 & data_did$group == 1)
#   indc_1 <- as.numeric(data_did$W == 0 & data_did$group == 1)
#   indt_0 <- as.numeric(data_did$W == 1 & data_did$group == 0)
#   indc_0 <- as.numeric(data_did$W == 0 & data_did$group == 0)
#   
#   Y_diff <- data_did$y_diff
#   X_mat <- prep$X_mat
#   
#   # Propensity scores
#   ps_1_t <- learning_function(Y = indt_1, X = X_mat, ids_out = ids, ids_in = ids, weights = data_did$weights)
#   ps_0_t <- learning_function(Y = indt_0, X = X_mat, ids_out = ids, ids_in = ids, weights = data_did$weights)
#   ps_1_c <- learning_function(Y = indc_1, X = X_mat, ids_out = ids, ids_in = ids, weights = data_did$weights)
#   ps_0_c <- learning_function(Y = indc_0, X = X_mat, ids_out = ids, ids_in = ids, weights = data_did$weights)
#   
#   # make sure non-zero
#   ps_0_t = pmax(ps_0_t, 1e-4)
#   ps_1_c = pmax(ps_1_c, 1e-4)
#   ps_0_c = pmax(ps_0_c, 1e-4)
#   
#   # Construct weights
#   # Construct weights
#   w_1_t       <- (indt_1) / mean(indt_1)
#   w_0_t       <- (indt_0) / mean(indt_0)
#   w_1_t_1_c   <- (indc_1 * ps_1_t / ps_1_c) / mean(indt_1)
#   w_1_t_0_t   <- (indt_0 * ps_1_t / ps_0_t) / mean(indt_1)
#   w_1_t_0_c   <- (indc_0 * ps_1_t / ps_0_c) / mean(indt_1)
#   w_0_t_0_c   <- (indc_0 * ps_0_t / ps_0_c) / mean(indt_1)
#   
#   # Calculate score functions
#   psi_ipw_1    <- (w_1_t - w_1_t_1_c) * (Y_diff)
#   psi_ipw_0    <- (w_0_t - w_0_t_0_c) * (Y_diff)
#   psi_wipw_1_0 <- (w_1_t_0_t - w_1_t_0_c) * Y_diff
#   psi_diff_dr <- psi_ipw_1 - psi_ipw_0
#   psi_tdid    <- psi_ipw_1 - psi_wipw_1_0
#   
#   # # Influence function adjustment for standard error estimation
#   eta.est         <- psi_tdid - mean(psi_tdid) * w_1_t
#   eta.diff_dr    <- psi_diff_dr - mean(psi_diff_dr) * w_1_t
#   
#   # Return a summary data frame of estimates
#   return(data.frame(
#     est              = mean(psi_tdid),
#     se.est           = sd(eta.est) / sqrt(length(eta.est)),
#     diff_dr          = mean(psi_diff_dr),
#     se.diff_dr       = sd(eta.diff_dr) / sqrt(length(eta.diff_dr)),
#     est_wdr_1_0      = mean(psi_wipw_1_0),
#     est_1            = mean(psi_ipw_1),
#     est_0            = mean(psi_ipw_0),
#     N                = length(psi_tdid)
#   ))
# }
```

Now can calculate DID by group and year

```{r}
# control_formula = "acc + alcohol + crack + ur"
# reference_year = 1993
# results = data.frame() 
# for(target_year in c(1986:1992, 1994:1996)) {
#     res = tdid_function(data_analysis, target_year, reference_year, control_formula) 
#     results = rbind(results, 
#                     res |> mutate(time = target_year))
# }
```

And plot this out

```{r}
# results |> 
#   ggplot(aes(x = time)) + 
#   geom_point(aes(y = est, color = "Correct TDID")) + 
#   geom_point(aes(y = diff_dr, color = "Wrong TDID")) + 
#   scale_x_continuous(breaks = 1986:1996)
# 
# results |> 
#   ggplot(aes(x = time)) + 
#   geom_point(aes(y = est_1, color = "DID Group 1 (Young)")) + 
#   geom_point(aes(y = est_0, color = "DID Group 0 (Old)")) + 
#   geom_point(aes(y = est_wdr_1_0, color = "Weighted DID Group 0 (Old)")) + 
#   scale_x_continuous(breaks = 1986:1996)
```

Using a similar function
